---
title: "NYPD Shooting Data"
author: "T. Pacheco"
date: '2022-07-25'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


Load Data
```{r NYPD Shooting Incidents}

url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"

library(tidyverse)

NYPD_shooting <- read_csv(url_in)


```





```{r Transformation}

NYPD_shooting <- subset(NYPD_shooting, select = -c(LOCATION_DESC, PERP_AGE_GROUP:PERP_RACE))

library(dplyr)
library(lubridate)

NYPD_shooting <- mutate(NYPD_shooting, OCCUR_DATE = mdy(OCCUR_DATE))

grouped_boro <- NYPD_shooting %>% 
  select(OCCUR_DATE, BORO) %>% 
  arrange(BORO)
```


Amount of shootings by Borough and shooting victims by gender

```{r Visualization}  

library(ggplot2)

ggplot(NYPD_shooting, aes(x = BORO)) + geom_bar() + xlab("Boroughs") + ylab("Shooting Incidents")

ggplot(NYPD_shooting, aes(x = "", y = VIC_SEX, fill = VIC_SEX)) + geom_col() + ylab("y") + coord_polar(theta = "y") + guides(fill = guide_legend(title = "Victim Sex")) + ggtitle("Incidents by Victim Sex") + scale_fill_discrete(labels = c("Female", "Male","Unkown"))


```


Amount of shootings per day reported in New York

The amount and percentage of shootings by borough

```{r Analyzation}
library(scales)

shootings_per_day = (length(NYPD_shooting$INCIDENT_KEY))/(length(unique(NYPD_shooting$OCCUR_DATE)))

cat("Shootings per day =",shootings_per_day)

total_shootings = length(NYPD_shooting$BORO)

incidents_by_boro <- grouped_boro %>%
  count(BORO) %>%
  rename(Shootings = n)

shooting_rate = (incidents_by_boro$Shootings/total_shootings)

percent_boro <- data.frame(incidents_by_boro, percent(shooting_rate))

percent_boro

incident_time = NYPD_shooting %>% 
  group_by(OCCUR_TIME) %>% 
  count()

ggplot(incident_time, aes(x = OCCUR_TIME, y = n)) + geom_line(color = "blue") + labs(title = "Time of Shootings", x = "Time", y = "Shootings")


```

As shown in the "Time of Shootings" Plot, the bulk of the shootings throughout New York take place when the sun is down. 

Based on the percentage of shootings the safest borough by gun violence is Staten Island.

Possible bias may come from the reporting of the shootings, the human element in data entry. 





